Landscape composition and beneficial insects

### 2020-05-15 manuscript/manuscript_outline.Rmd
Worked on Methods/Results

#Adds ** around text (makes it bold in RMarkdown)
boldRMD <- function(x) paste0('**',x,'**')

@@

@@

* To characterize landscape composition surrounding the traps, we used classified landscape data from AAFC
  + Classifications are very accurate for fields (~90%), but less accurate for non-crop areas (~70%, [@AAFC_data])
  + Landscape studies often use the landscape composition within a circle of a fixed radius surrounding each point
  + However, the radius of the circle depends on the grain of the landscape that is relevant to the abundance of a given organism.
  + Instead, we divided the landscape into annuli (rings) surrounding each trap location.
  + Each annulus was 30 m thick, and had inner radii ranging from 30 to 1470m, in increments of 30m.
  + Using functional regression (see below), we were able to incorporate landscape information at different distance, allowing us to assess the impact of both local and regional landscape composition

@@

* Function-on-scalar (FOS) regression is a special type of linear model ($\hat{y}=X\beta$), where the values of the coefficients ($\beta$) are defined by a smooth function
  + In our case, the coefficients represent the effect of a given type of landscape cover, and are a smooth function of distance from the trap (i.e. $\beta_i = f(distance_i)$)
  + This technique has been used by other authors; @galpern2020 called this ``onion-skin regression"
  + We fit our functional regression models using _mgcv_ version 1.8-31.

@@

* _P. melanarius_ activity density was strongly influenced by trap location and landscape composition (Tables \@ref(tab:PteMelCoefs-linTerms), \@ref(tab:PteMelCoefs-smoothTerms))
  + Canola crops had a higher activity density of _P. melanarius_ than any other trap location, while roadside ditches had the lowest (Figure \@ref(fig:PteMelCoefs-fixeff)a)
  + This indicates that low-density stands of canola may allow faster movement of ground beetles.
  + In landscape composition, proportion of urban land cover (roadside right-of-way) was the strongest predictor of activity density, and this varied across the season, with urban land acting as a sink during June, but becoming a source during July (Figure \@ref(fig:PteMelCoefs-fixeff)f); trees and shrubs followed a similar pattern (Figure \@ref(fig:PteMelCoefs-fixeff)d)
  + Wetland cover had a negative influence on _P. melanarius_ abundance, but only after mid-July, indicating that wetlands may act as a sink for _P. melanarius_ populations later in the season

```{r PteMelCoefs-linTerms, echo=FALSE, warning=FALSE}
pmelRes <- summary(PteMelMod$mod3)
linTerms <- pmelRes$p.table %>% as.data.frame() %>% 
  rownames_to_column('loc') %>% 
  mutate(loc=firstUpper(gsub('trapLoc','',loc))) %>% 
  rename(p='Pr(>|z|)') %>% mutate(p=roundLess(p,3)) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
kable(linTerms,caption='\\label{tab:PteMelCoefs-linTerms} Linear terms for _Pterostichus melanarius_. Line',col.names=c('Trap location','$\\beta$','S.E.','Z','p'),
        digits=c(1,2,2,2,2),align='rrrrl')
```

```{r PteMelCoefs-smoothTerms, echo=FALSE, warning=FALSE}
pmelRes <- summary(PteMelMod$mod3)
smTerms <- pmelRes$s.table %>% as.data.frame() %>% 
  rownames_to_column('Smoother') %>% 
  rename(p='p-value') %>% mutate(p=roundLess(p,3)) %>% mutate_at(vars(2:4),roundLess,dig=2) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
kable(smTerms[c(1,2,4,5)],caption='\\label{tab:PteMelCoefs-smoothTerms} Smooth terms for _Pterostichus melanarius_, with effective degrees of freedom, Chi-squared coefficient, and p-value. _s_ indicates a thin-plate spline, _ti_ indicates a tensor-product interaction. All terms except for day and (E,N) are functional regression fits.',col.names=c('Smoothing term','Eff. _df_','$\\chi^2$','_p_'),
        digits=c(1,2,2,2),align='rrll')
```

@@

*  _P. melanarius_ had a very strong random temporal and spatial component (both p<0.001), indicating that phenology and local geographic factors were strong drivers of ground beetle abundance (Figure \@ref(fig:PteMelCoefs-raneff))
  + This could be related to climatic drivers (rainfall, growing degree days), or could simply represent a single yearly realization of a spatio-temporal random field

@@

```{r ParDisCoefs-linTerms, echo=FALSE, warning=FALSE}
pmelRes <- summary(ParDisMod$mod3)
linTerms <- pmelRes$p.table %>% as.data.frame() %>% 
  rownames_to_column('loc') %>% 
  mutate(loc=firstUpper(gsub('trapLoc','',loc))) %>% 
  rename(p='Pr(>|z|)') %>% mutate(p=roundLess(p,3)) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
kable(linTerms,caption='\\label{tab:ParDisCoefs-linTerms} Linear terms for _Pardosa distincta_. Line',col.names=c('Trap location','$\\beta$','S.E.','Z','p'),
        digits=c(1,2,2,2,2),align='rrrrl')
```

```{r ParDisCoefs-smoothTerms, echo=FALSE, warning=FALSE}
pmelRes <- summary(ParDisMod$mod3)
smTerms <- pmelRes$s.table %>% as.data.frame() %>% 
  rownames_to_column('Smoother') %>% 
  rename(p='p-value') %>% mutate(p=roundLess(p,3)) %>% mutate_at(vars(2:4),roundLess,dig=2) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
kable(smTerms[c(1,2,4,5)],caption='\\label{tab:ParDisCoefs-smoothTerms} Smooth terms for _Pardosa distincta_, with effective degrees of freedom, Chi-squared coefficient, and p-value. _s_ indicates a thin-plate spline, _ti_ indicates a tensor-product interaction. All terms except for day and (E,N) are functional regression fits.',col.names=c('Smoothing term','Eff. _df_','$\\chi^2$','_p_'),
        digits=c(1,2,2,2),align='rrll')
```

@@

```{r OpilioCoefs-linTerms, echo=FALSE, warning=FALSE}
pmelRes <- summary(OpilioMod$mod3)
linTerms <- pmelRes$p.table %>% as.data.frame() %>% 
  rownames_to_column('loc') %>% 
  mutate(loc=firstUpper(gsub('trapLoc','',loc))) %>% 
  rename(p='Pr(>|z|)') %>% mutate(p=roundLess(p,3)) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
kable(linTerms,caption='\\label{tab:OpilioCoefs-linTerms} Linear terms for _Phalangium opilio_. Line',col.names=c('Trap location','$\\beta$','S.E.','Z','p'),
        digits=c(1,2,2,2,2),align='rrrrl')
```

```{r OpilioCoefs-smoothTerms, echo=FALSE, warning=FALSE}
pmelRes <- summary(OpilioMod$mod3)
smTerms <- pmelRes$s.table %>% as.data.frame() %>% 
  rownames_to_column('Smoother') %>% 
  rename(p='p-value') %>% mutate(p=roundLess(p,3)) %>% mutate_at(vars(2:4),roundLess,dig=2) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
kable(smTerms[c(1,2,4,5)],caption='\\label{tab:OpilioCoefs-smoothTerms} Smooth terms for _Phalangium opilio_, with effective degrees of freedom, Chi-squared coefficient, and p-value. _s_ indicates a thin-plate spline, _ti_ indicates a tensor-product interaction. All terms except for day and (E,N) are functional regression fits.',col.names=c('Smoothing term','Eff. _df_','$\\chi^2$','_p_'),
        digits=c(1,2,2,2),align='rrll')
```

@@

```{r OpilioCoefs-raneff, echo=FALSE, fig.cap='\\label{fig:OpilioCoefs-raneff} Temporal and spatial components of _Phalangium opilio_ activity density'}
  include_graphics('../figures/Opiliones_raneff.png')

@@

>>>>

#Adds ** around text (makes it bold in RMarkdown)
boldRMD <- function(x) paste0('**',x,'**')
#Inverse of above
unboldRMD <- function(x) gsub('\\**','',x)

@@

#Extract Pterostichus melanarius results
pmelRes <- summary(PteMelMod$mod3)
pmelLinTerms <- pmelRes$p.table %>% as.data.frame() %>% 
  rownames_to_column('loc') %>% 
  mutate(loc=firstUpper(gsub('trapLoc','',loc))) %>% 
  rename(p='Pr(>|z|)') %>% mutate(p=roundLess(p,3)) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
pmelSmTerms <- pmelRes$s.table %>% as.data.frame() %>% 
  rownames_to_column('Smoother') %>% 
  rename(p='p-value') %>% mutate(p=roundLess(p,3)) %>% mutate_at(vars(2:4),roundLess,dig=2) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold

#Extract Pardosa distincta results
pdisRes <- summary(ParDisMod$mod3)
pdisLinTerms <- pdisRes$p.table %>% as.data.frame() %>% 
  rownames_to_column('loc') %>% 
  mutate(loc=firstUpper(gsub('trapLoc','',loc))) %>% 
  rename(p='Pr(>|z|)') %>% mutate(p=roundLess(p,3)) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
pdisSmTerms <- pdisRes$s.table %>% as.data.frame() %>% 
  rownames_to_column('Smoother') %>% 
  rename(p='p-value') %>% mutate(p=roundLess(p,3)) %>% mutate_at(vars(2:4),roundLess,dig=2) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold

#Extract Phalangium opilio results
opilRes <- summary(OpilioMod$mod3)
opilLinTerms <- opilRes$p.table %>% as.data.frame() %>% 
  rownames_to_column('loc') %>% 
  mutate(loc=firstUpper(gsub('trapLoc','',loc))) %>% 
  rename(p='Pr(>|z|)') %>% mutate(p=roundLess(p,3)) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold
opilSmTerms <- opilRes$s.table %>% as.data.frame() %>% 
  rownames_to_column('Smoother') %>% 
  rename(p='p-value') %>% mutate(p=roundLess(p,3)) %>% mutate_at(vars(2:4),roundLess,dig=2) %>% 
  mutate(sig=is.na(as.numeric(p))|as.numeric(p)<0.05) %>% #If p<0.05
  mutate(p=ifelse(sig,boldRMD(p),p)) %>% select(-sig) #Make text bold

@@

* To characterize landscape composition surrounding the traps, we used classified landscape data from AAFC
  + Classifications are very accurate for fields (~90%), but less accurate for non-crop areas (~70%, [@AAFC_data])
  + Landscape studies often use the landscape composition within a circle of a fixed radius surrounding each point
  + However, the radius of the circle depends on the grain of the landscape that is relevant to the abundance of a given organism (@addicott1987)
  + Instead, we divided the landscape into annuli (rings) surrounding each trap location.
  + Each annulus was 30 m thick, and had inner radii ranging from 30 to 1470m, in increments of 30m.
  + Using functional regression (see below), we were able to incorporate landscape information at different distance, allowing us to assess the impact of both local and regional landscape composition

@@

* Scalar-on-function (SoF) regression is a special type of linear model ($\hat{y}=X\beta$), where the columns of the model matrix $X$ contain some continuous predictor of the scalar $y$ (e.g. biomass as a function of monthly average temperatures), and the values of the coefficients ($\beta$) are modeled as a smooth function $f(x)$ of the continuous predictors
  + In our case, each column represents the landscape composition within each annulus, with increasing radii away from the location of the trap, and the coefficients are a smooth function of distance ($\beta_i = f(distance_i)$), meaning that the coefficients represent the (additive) effect of a given type of landscape cover and  from the trap 
  + This avoids the fundamental problem of choosing an optimal radius of landscape composition, and allows the possibility that the size of the landscape ``grain" relevant to a given organism may change over the course of its life stages (@addicott1987, @lima1996)
  + This technique has been used by other authors (``onion-skin regression" in @galpern2020)
  + We fit our functional regression models using _mgcv_ version 1.8-31.

<!-- This part could be pared down. Probably not necessary to discuss several models if the full one is the only important one. -->
* We used 4 nested models to gauge the relative importance of landscape composition for arthropod activity density
  + Model 1 consisted of a temporal and spatial smoothers (``random effects"), necessary to account for the underlying spatiotemporal structure of the data 
  <!-- Smoothers can be thought of as random effects, where the coefficients follow a probability distribution defined by a fixed smoothing coefficient $\lambda$ - see @wood2017 for the Bayesian/Frequentist details of this -->
  + Model 2 added a local cover term, to account for the specific effect of local cover on activity density
  + Model 3 added a smoother term for the landscape rings, modeling the effect of non-crop cover at varying distances on activity density
  + Model 4, similar to model 3, modeled the effect of all underlying cover classes (crop and non-crop) at varying distances
  + To account for different lengths of trap exposure, log-days since trap placement was used as a fixed effect with its slope held at 1 (``offset" variable)
  + All models were fit using a negative binomial distribution, and model fits were compared using conditional Akaike's Information Criteria @wood2017

Model    | Independent variables                                     | $r$ Formula
-------- | --------------------------------------------------------- | -----------------------------------------------
1        | Phenology + Geography                                     | Activity Density ~ s(Day) + s(E,N)
2        | Phenology + Geography + Local cover                       | Activity Density ~ s(Day) + s(E,N) + Trap Location
3        | Phenology + Geography + Local cover + Landscape (general) | Activity Density ~ s(Day) + s(E,N) + Trap Location + s(Non-crop Cover)                         
4        | Phenology + Geography + Local cover + Landscape (specific)| Activity Density ~ s(Day) + s(E,N) + Trap Location + s(Cover 1) + s(Cover 2) + ... + s(Cover n)

@@

* _P. melanarius_ activity density was strongly influenced by trap location and landscape composition (Tables \@ref(tab:PteMelCoefs-linTerms), \@ref(tab:PteMelCoefs-smoothTerms))
  + Canola crops had a higher activity density of _P. melanarius_ than any other trap location, while roadside ditches had the lowest (Figure \@ref(fig:PteMelCoefs-fixeff)a)
  + This indicates that low-density stands of canola may allow faster movement of ground beetles.
  + In landscape composition, proportion of urban land cover (roadside right-of-way) was the strongest predictor of activity density, and this varied across the season, with urban land acting as a sink during June, but becoming a source during July (Figure \@ref(fig:PteMelCoefs-fixeff)f); trees and shrubs followed a similar pattern (Figure \@ref(fig:PteMelCoefs-fixeff)d)
  + Wetland cover had a negative influence on _P. melanarius_ activity density, but only after mid-July, indicating that wetlands may act as a sink for _P. melanarius_ populations later in the season

```{r PteMelCoefs-linTerms, echo=FALSE, warning=FALSE}
kable(pmelLinTerms,caption='\\label{tab:PteMelCoefs-linTerms} Linear terms for _Pterostichus melanarius_.',col.names=c('Trap location','$\\beta$','S.E.','Z','p'),
        digits=c(1,2,2,2,2),align='rrrrl')
```

```{r PteMelCoefs-smoothTerms, echo=FALSE, warning=FALSE}
kable(pmelSmTerms[c(1,2,4,5)],caption='\\label{tab:PteMelCoefs-smoothTerms} Smooth terms for _Pterostichus melanarius_, with effective degrees of freedom, Chi-squared coefficient, and p-value. _s_ indicates a thin-plate spline, _ti_ indicates a tensor-product interaction. All terms except for day and (E,N) are functional regression fits. **NAMES OF SMOOTHING TERMS ARE CONFUSING - SHOULD BE SIMPLIFIED**',col.names=c('Smoothing term','Eff. _df_','$\\chi^2$','_p_'),
        digits=c(1,2,2,2),align='rrll')
```

@@

*  _P. melanarius_ had a very strong random temporal and spatial component (both p<0.001), indicating that phenology and local geographic factors were strong drivers of ground beetle activity density (Figure \@ref(fig:PteMelCoefs-raneff))
  + This could be related to climatic drivers (rainfall, growing degree days), or could simply represent a single yearly realization of a spatio-temporal random field

@@

* _Pardosa distincta_ activity density was also strongly influenced by trap location and landscape composition (Tables \@ref(tab:ParDisCoefs-linTerms), \@ref(tab:ParDisCoefs-smoothTerms)), similar to _P. melanarius_.
  + However, _P. distincta_ had far lower activity density in canola than any other cover type (Figure \@ref(fig:ParDisCoefs-fixeff)a).
  + Interestingly, nearby canola appeared to act as a source of _P. distincta_ for surrounding landscape features, as activity density was positively related to the nearby proportion of canola (p = `r with(pdisSmTerms,unboldRMD(p[Smoother=='s(distMat):Canola']))`), at least early in the season (p = `r with(pdisSmTerms,unboldRMD(p[Smoother=='s(endDayMat):Canola']))`)
  + Pulse crops had a similar effect, acting as a source of _P. distincta_ early in the season, and acting as a mild sink later in the season (p = `r with(pdisSmTerms,unboldRMD(p[Smoother=='ti(distMat,endDayMat):Pulses']))`, Figure \@ref(fig:ParDisCoefs-fixeff)d)
  + Grassland acted as a weak source, at least earlier on in the season (p = `r with(pdisSmTerms,unboldRMD(p[Smoother=='s(endDayMat):Grassland']))`), and urban cover acted as a weak sink at intermediate distances (~600m)
  + Similar to _P. melanarius_, _P. distincta_ activity density had a very strong temporal and spatial component (both p = `r with(pdisSmTerms,unboldRMD(p[grepl('E,N',Smoother)]))`, Figure \@ref(fig:ParDisCoefs-raneff))

```{r ParDisCoefs-linTerms, echo=FALSE, warning=FALSE}
kable(pdisLinTerms,caption='\\label{tab:ParDisCoefs-linTerms} Linear terms for _Pardosa distincta_. Line',col.names=c('Trap location','$\\beta$','S.E.','Z','p'),
        digits=c(1,2,2,2,2),align='rrrrl')
```

```{r ParDisCoefs-smoothTerms, echo=FALSE, warning=FALSE}
kable(pdisSmTerms[c(1,2,4,5)],caption='\\label{tab:ParDisCoefs-smoothTerms} Smooth terms for _Pardosa distincta_, with effective degrees of freedom, Chi-squared coefficient, and p-value. _s_ indicates a thin-plate spline, _ti_ indicates a tensor-product interaction. All terms except for day and (E,N) are functional regression fits.',col.names=c('Smoothing term','Eff. _df_','$\\chi^2$','_p_'),
        digits=c(1,2,2,2),align='rrll')
```

@@

* _Phalangium opilio_ activity density was strongly influenced by trap location, with pivot corners and wetlands having the highest activity density (p = `r with(opilLinTerms, unboldRMD(p[loc=='Pivot']))`, `r with(opilLinTerms, unboldRMD(p[loc=='Wetland']))`)  
  + However, unlike _P. melanarius_ or _P. pardosa_, _P. opilio_ was not strongly influenced by any land cover type, except for Pasture (p = `r with(opilSmTerms,unboldRMD(p[Smoother=='ti(distMat,endDayMat):Pasture']))`)
  + This specific effect of pasture was not large, as the model using non-crop cover performed similarly ($\Delta AIC$= `r with(OpilioMod, sprintf("%.2f",AIC(mod3)-AIC(mod4)))`)
  + In fact, both models that used landscape composition performed worse than the model with only trap location ($\Delta AIC$= `r with(OpilioMod, sprintf("%.2f",AIC(mod2)-AIC(mod3)))`,`r with(OpilioMod, sprintf("%.2f",AIC(mod2)-AIC(mod4)))`)

```{r OpilioCoefs-linTerms, echo=FALSE, warning=FALSE}
kable(opilLinTerms,caption='\\label{tab:OpilioCoefs-linTerms} Linear terms for _Phalangium opilio_. Line',col.names=c('Trap location','$\\beta$','S.E.','Z','p'),
        digits=c(1,2,2,2,2),align='rrrrl')
```

```{r OpilioCoefs-smoothTerms, echo=FALSE, warning=FALSE}

kable(opilSmTerms[c(1,2,4,5)],caption='\\label{tab:OpilioCoefs-smoothTerms} Smooth terms for _Phalangium opilio_, with effective degrees of freedom, Chi-squared coefficient, and p-value. _s_ indicates a thin-plate spline, _ti_ indicates a tensor-product interaction. All terms except for day and (E,N) are functional regression fits.',col.names=c('Smoothing term','Eff. _df_','$\\chi^2$','_p_'),
        digits=c(1,2,2,2),align='rrll')
```

@@

```{r OpilioCoefs-raneff, echo=FALSE, fig.cap='\\label{fig:OpilioCoefs-raneff} Temporal and spatial components of _Phalangium opilio_ activity density'}
  include_graphics('../figures/Opiliones_raneff.png')
```

@@

* Lemme tell ya 
  + We found all kinds of stuff
  + Boy oh boy is it cool
